---
title: "hw2"
author: "Ritika Agrawal"
github username: "ritikagrawal1710"
date: "February 7, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This homework is about syntax and basic plotting functions in R, and is **due by the end of Feb 7**. 

You can use the RMarkDown file with these homework instructions as a template for your work. 
Make sure to replace "Your Name" and "your_github_username" in the YAML with your name and Github username.

**Submission instructions**: Create a folder named `ds202_hw2`, and name the RMarkDown file including your solutions `hw2.Rmd` under this folder. For submission, create a GitHub repository named `ds202_hw2` under your account, and push both `hw2.Rmd` and the knitted `hw2.html` to GitHub before the deadline. I will assume you use the same GitHub username as for your HW1 submission. The latest version of your homework appearing on GitHub before the deadline is graded. *It is crucial to follow the exact folder structure and file names*, so that your homework can be reproduced and graded by an automated script. Some questions will require the following answer format.

## Example question and answer
0\. What is the smallest positive odd integer? Storage it in a variable named `a0` and print it.

**Example answer**:
In the RMarkDown file, type in
<pre><code>```{r}
a0 <- 1
a0
```</code></pre>
The knitted html file will show:
```{r}
a0 <- 1
a0
```

## R commands
1. Today's temperature has a high of 24 and a low of 15 degree fahrenheit.  Find the corresponding degress in celsius and storage it in a (single) vector named `cel`. Print the variable.
```{r}
cel <- c(((24-32)*5/9),((15-32)*5/9))
print(cel)
```

2. Create a vector of length 200. The odd entries should be 3 and even entries should be 1. Storage the created in a variable named `vec200` and print.
```{r}
n=200

vec200 <- c()
for (i in 1:n){
  
  if (i%%2){
    
    vec200 <- append(vec200,3)
  }
    else{ 
      vec200 <- append(vec200,1)
  }
}
print (vec200)
```

3. How many perfect squares are there between 1 and 2001? Storage the answer in a  variable named `numPerfect` and print. You may want to look at `?round`.
```{r}
n= 2001
numperfect = 0
for (i in 1:n){
  s = sqrt(i)
  r = round(s)
  if (s==r){
    numperfect <- numperfect+1
  }
}
print(numperfect)
```

## Cars and mileage
For this part we will use the `mpg` data set, you can access the data by first installing (using `install.packages('ggplot2')`) and then loading the package `ggplot2` (using `library(ggplot2)`) into your R session. The `mpg` data includes information about fuel economy of popular car models
for the years 1999 and 2008 as collected by the US Environmental Protection Agency. An answer should include a short paragraph of text for the explanation to a (human) reader, together with appropriate R code, outputs, and graphs.


install.packages("ggplot2")
```{r}
library(ggplot2)
head(mpg)
```

Most exercises are from the ggplot2 book section 2.2.1. You might find it helpful for working on the homework to read through this section. *However, all questions can be solved by using just the basic R graphics functions.*

1. What are the 3 most fuel-efficient models in terms of highway mileage? Storage the data frame (with 3 rows) containing all information of these three cars in a variable named `top3` and print.

```{r}

df <- mpg[order(mpg$hwy,decreasing = TRUE),]
top3 <- df[1:3,]
top3
```
2. How many different compact models are included in this dataset? Storage the answer in  a variable named `numCompact` and print.
```{r}
df_compact_Models <- mpg[mpg$class=="compact",]
numCompact <- dim(df_compact_Models)[1]
numCompact
```

3. Create a scatterplot of `hwy` and `cty` and describe the relationship. Why are there so few points visible? Explain your findings.
```{r}
plot(mpg$hwy, mpg$cty, main="Scatterplot Example", xlab="hwy", ylab="cty", pch=19)
```

Response: The relationship between 'hwy' and 'cty' is linear. As 'hwy' increases, 'cty' also increases. 
The values are overlapped, hence only few points are visible. 

4. Are cars in 2008 more fuel efficient than cars in 1999? Use graphs *and* summary statistics to answer this question. Explain your findings.
```{r}
boxplot(cty~year,data=mpg, main="Car Milage Data",
        xlab="year", ylab="Miles Per Gallon")

boxplot(hwy~year,data=mpg, main="Car Milage Data",
        xlab="year", ylab="Miles Per Gallon")

mpg_2008 = mpg[mpg$year==2008,]
mpg_1999 = mpg[mpg$year==1999,]

head(mpg_2008)
head(mpg_1999)

summary(mpg_2008)
summary(mpg_1999)
```
Response: I have drawn box-plots for both 'cty' and 'hwy' mileage. As can be seen from the box-plot for 'cty', the median appears to be the same indicating that fuel efficienct of cars for both years 1999 and 2008 is the same. However, in the box-plot of the 'hwy' mileage, it is clearly visible that the efficiency of cars for year 2008 is better than that of year 1999. 
When we see the summary for the two years, the same output is visbile. The median for 'cty' mileage in year 1999 and year 2008 is 17.0 just as we saw in the box-plot. In the summary for 'hwy' mileage, the median for year 1999 is 25 whereas for year 2008 it is 24, thus indicating better efficiency in year 2008, just as we saw in the box-plot.

Note: your submission is supposed to be fully reproducible, i.e. the TA and I will 'knit' your submission. Make sure that you don't use any file structure that depends on your computing environment.